Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-10 fib2.tex
edits: looking good to me

From the perspective of
\emph{topological quantum field theory}
(TQFT), anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid, but the manifold in which the anyons embed 
is considered deformed (twisted) by these braid moves.
%The generators corresponding to the braid
%group generators are called half-twists.
% -we thought this was a getting ahead of ourselves since we haven't introduced this stuff yet.
This is the perspective we adopt.
In particular, this shows the action of braiding
on arbitrary observables (Wilson loops) as
these deform with the manifold.

@@

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways.
\dude{picture? ref Fig 2?}
%
%
%Braiding within an observable will only change the phase
%of that observable \dude{???}, otherwise may twist these
%loops in various ways:
\begin{align*}
\includegraphics[]{pic-braid-loop.pdf}
\end{align*}
%
%
These diagrams can be composed, and the fusion operation replaces the 
interior of a disc by a single charge.
These pictures, abstracted from the quantum context,
are called \emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
%In general, such a curve diagram is 
%a continuous embedding of a line that passes through 
%each puncture once. 
%Remarkably, any two such linear orderings are related by
%braiding (half-twists).
%\dude{optional paragraph..}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ observables arranged in
a square lattice tiling $\Lambda$.
These \emph{tiles} will be the oservables accesible to the error-correction procedure (\emph{decoder}).
We imagine an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\vac}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_i$ the projector to charge $\vac$ at tile $i$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
\dude{ouch this no longer works..}
Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above.
We identify this ground space as the codespace of our model. 
Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

At some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, we use heuristics
to minimize non-local braid moves, which helps to
maintain the state as a separable product. \dude{???}
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
%However, for finite lattice size, this regime is somewhere
%beyond the decoder threshold, which is the 
%point at which the decoder creates nontrivial (extensive) loops.
%\stf{CGB and I are still not confident in the previous sentence or the next one}
%We can view the system state as a tensor network, on which
%the decoder is performing an (efficiently simulable) contraction.
%This enables simulation of error correction in
%the area around decoder threshold for lattice sizes up to $L=128.$
%\dude{more polish needed...}

>>>>

From the perspective of
\emph{topological quantum field theory}
anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid, but the manifold in which the anyons reside 
becomes deformed (twisted) by these braid moves.
This is the perspective we adopt.
In particular, this shows the action of braiding
on arbitrary observables (Wilson loops) as
these deform with the manifold.

@@

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways:
\begin{align*}
\includegraphics[]{pic-braid-loop.pdf}
\end{align*}
%
%
These diagrams can be composed, and the fusion operation replaces the 
interior of a disc by a single charge.
These pictures, abstracted from the quantum context,
are called \emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
%In general, such a curve diagram is 
%a continuous embedding of a line that passes through 
%each puncture once. 
%Remarkably, any two such linear orderings are related by
%braiding (half-twists).
%\dude{optional paragraph..}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ observables arranged in
a square lattice tiling $\Lambda$.
These \emph{tiles} will be the oservables accesible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{T\in \Lambda}\proj{\vac}_T\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_T$ the projector to charge $\vac$ at tile $T$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space contains the two-fold degenerate codespace.
%We identify this ground space as the codespace of our model. 
Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

At some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, we use heuristics
to minimize non-local braid moves, which helps to
maintain the state as a separable product.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
%However, for finite lattice size, this regime is somewhere
%beyond the decoder threshold, which is the 
%point at which the decoder creates nontrivial (extensive) loops.
%\stf{CGB and I are still not confident in the previous sentence or the next one}
%We can view the system state as a tensor network, on which
%the decoder is performing an (efficiently simulable) contraction.
%This enables simulation of error correction in
%the area around decoder threshold for lattice sizes up to $L=128.$
%\dude{more polish needed...}
